### Tidying our data for analysis

```{r Installing packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(ggplot2)
library(scales)
library(grid)
library(gridExtra)
```

###Import the required datasets

```{r Loading databases, message=FALSE, warning=FALSE}
movie_df_1 <- read_csv("../01_raw_data_sets/MoviesOnStreamingPlatforms.csv")
tv_df_1 <- read_csv("../01_raw_data_sets/tv_shows.csv")
```

###Clean data to prep for data manipulation and transformation
#### Tidying our dataframes
##### Movie data frame

We will start by cleaning the names, then selecting the columns we're interested in examining. Then we'll make each column name more distinct to clarify when we're analyzing both data frames.
```{r}
clean_movie_df_1 <- clean_names(movie_df_1)
movie_df_2 <- clean_movie_df_1 %>% 
  select("id", "year", "age", "rotten_tomatoes", "netflix", "hulu", "prime_video", "disney")
movie_df_2 <- movie_df_2 %>% 
  rename(release_year_movie=year, age_rating_movie=age, rotten_tomatoes_movie=rotten_tomatoes, netflix_movie=netflix, hulu_movie=hulu, prime_video_movie=prime_video, disney_movie=disney)
glimpse(movie_df_2)
```
After a quick look at the movie data frame we've put together, we see it has 9,515 rows and the 8 columns we selected.  Now we resolve any NA's in our data frame, but first we check for them.
```{r}
sum(is.na(movie_df_2))
sum(is.na(movie_df_2$age_rating_movie))
sum(is.na(movie_df_2$rotten_tomatoes_movie))
```
It looks like age_rating_movie has most of the NA's (4,177), and rotten_tomatoes_movie as the rest (7).  We have a lot of Na's in age_rating_movie, so we're going to replace these rather than remove them. For rotten_tomatoes_movie, there are only 7 NA's in over 9500 rows, so we will remove them without skewing data.
```{r}
movie_df_2 <- replace_na(movie_df_2, list(age_rating_movie = "unknown"))
movie_df_2 <- drop_na(movie_df_2)
```

##### TV data frame
Next, we'll follow the same procedure for the TV data set, beginning with cleaning the names of each column and then selecting columns for a new data frame. We will also rename the columns for clarity. 
```{r}
clean_tv_df_1 <- clean_names(tv_df_1)
tv_df_2 <- clean_tv_df_1 %>% 
  select("id", "year", "age", "rotten_tomatoes", "netflix", "hulu", "prime_video", "disney")
tv_df_2 <- tv_df_2 %>% 
  rename(release_year_tv=year, age_rating_tv=age, rotten_tomatoes_tv=rotten_tomatoes, netflix_tv=netflix, hulu_tv=hulu, prime_video_tv=prime_video, disney_tv=disney)
glimpse(tv_df_2)
```

The TV data frame is smaller than the Movie data frame, with 5,368 rows and 8 columns. Now we're going to check for NA's. 
```{r}
sum(is.na(tv_df_2))
sum(is.na(tv_df_2$age_rating_tv))
```

All of the NA's are in the age category. There are too many to remove without skewing data, so we'll replace. 
```{r}
tv_df_2 <- replace_na(tv_df_2, list(age_rating_tv = "unknown"))
sum(is.na(tv_df_2))
```

#### Rotten tomatoes conversions
The next step to prepping our data is to convert the rotten_tomatoes for both data frames from character to number
This will be done is a couple steps. First, we'll separate the value, so it does not read as a fraction when we convert, then convert from a character type vector to a double. We'll do this for both data frames. 

##### Movie data frame
```{r Movie df rotten tomatoes to double, message=FALSE, results='hide'}
movie_df_2 <- separate(movie_df_2, rotten_tomatoes_movie, sep="/", into=c("rotten_tomatoes_movie", "soon_to_be_dropped")) %>% 
select(-soon_to_be_dropped)
sapply(movie_df_2, class)
vec <- c(4)
movie_df_2[ , vec] <- apply(movie_df_2[ , vec, drop=F], 1,
                                                  function(x) as.numeric(as.character(x)))
```

##### TV data frame
We'll repeat the same separation and conversion with the TV data frame.
```{r TV df rotten tomatoes to double, results='hide'}
tv_df_2 <- separate(tv_df_2, rotten_tomatoes_tv, sep="/", into=c("rotten_tomatoes_tv", "soon_to_be_dropped_2")) %>% 
  select(-soon_to_be_dropped_2)
sapply(tv_df_2, class)
vec <- c(4)
tv_df_2[ , vec] <- apply(tv_df_2[ , vec, drop=F], 1,
                                               function(x) as.numeric(as.character(x)))
```

#### Collective platform column
There is no column containing all the streaming platforms, so we will make one now, as a new data frame, since we are making several vector type changes.First, we copy our data frame, then name changes from double vectors to characters. Then remane them, so that they remain distinct with we put them together in one column. 

##### Movie data frame
```{r movie platforms from double to character, results='hide'}
movie_df_3 <- movie_df_2
glimpse(movie_df_3)

sapply(movie_df_3, class)
vec <- c(5)
movie_df_3[ , vec] <- apply(movie_df_3[ , vec, drop=F], 1,
                                          function(x) as.character(as.numeric(x)))
sapply(movie_df_3, class)
vec <- c(6)
movie_df_3[ , vec] <- apply(movie_df_3[ , vec, drop=F], 1,
                                          function(x) as.character(as.numeric(x)))
sapply(movie_df_3, class)
vec <- c(7)
movie_df_3[ , vec] <- apply(movie_df_3[ , vec, drop=F], 1,
                                          function(x) as.character(as.numeric(x)))
sapply(movie_df_3, class)
vec <- c(8)
movie_df_3[ , vec] <- apply(movie_df_3[ , vec, drop=F], 1,
                                          function(x) as.character(as.numeric(x)))
```

We have now converted all the streaming platforms columns to character vectors. We'll substitute the "1's" and "0's" as longer, distinct strings, trim white space, and make our new column.

```{r movie platforms 1 to string, results='hide'}
movie_df_3$netflix_movie <- gsub("1", "netflix_film", movie_df_3$netflix_movie)
movie_df_3$netflix_movie <- gsub("0", "", movie_df_3$netflix_movie)

movie_df_3$hulu_movie <- gsub("1", "hulu_film", movie_df_3$hulu_movie)
movie_df_3$hulu_movie <- gsub("0", "", movie_df_3$hulu_movie)

movie_df_3$prime_video_movie <- gsub("1", "prime_video_film", movie_df_3$prime_video_movie)
movie_df_3$prime_video_movie <- gsub("0", "", movie_df_3$prime_video_movie)

movie_df_3$disney_movie <- gsub("1", "disney_film", movie_df_3$disney_movie)
movie_df_3$disney_movie <- gsub("0", "", movie_df_3$disney_movie)

movie_df_3$all_platforms_movie <- paste(movie_df_3$netflix_movie , movie_df_3$hulu_movie , movie_df_3$prime_video_movie, movie_df_3$disney_movie)
movie_df_3$all_platforms_movie <- trimws(movie_df_3$all_platforms_movie , which = c("right"))
movie_df_3$all_platforms_movie <- 
  ifelse((nchar(movie_df_3$all_platforms_movie)) <= 20, movie_df_3$all_platforms_movie, "multi_platform_film" )

```

##### TV data frame
We repeat this conversion for the TV data frame.
```{r movie platforms from double to character and 1 to string, results='hide'}
tv_df_3 <- tv_df_2
sapply(tv_df_3, class)
vec <- c(5)
tv_df_3[ , vec] <- apply(tv_df_3[ , vec, drop=F], 1,
                           function(x) as.character(as.numeric(x)))
sapply(tv_df_3, class)
vec <- c(6)
tv_df_3[ , vec] <- apply(tv_df_3[ , vec, drop=F], 1,
                           function(x) as.character(as.numeric(x)))
sapply(tv_df_3, class)
vec <- c(7)
tv_df_3[ , vec] <- apply(tv_df_3[ , vec, drop=F], 1,
                           function(x) as.character(as.numeric(x)))
sapply(tv_df_3, class)
vec <- c(8)
tv_df_3[ , vec] <- apply(tv_df_3[ , vec, drop=F], 1,
                           function(x) as.character(as.numeric(x)))
tv_df_3$netflix_tv <- gsub("1", "netflix_show", tv_df_3$netflix_tv)
tv_df_3$netflix_tv <- gsub("0", "", tv_df_3$netflix_tv)

tv_df_3$hulu_tv <- gsub("1", "hulu_show", tv_df_3$hulu_tv)
tv_df_3$hulu_tv <- gsub("0", "", tv_df_3$hulu_tv)

tv_df_3$prime_video_tv <- gsub("1", "prime_video_show", tv_df_3$prime_video_tv)
tv_df_3$prime_video_tv <- gsub("0", "", tv_df_3$prime_video_tv)

tv_df_3$disney_tv <- gsub("1", "disney_show", tv_df_3$disney_tv)
tv_df_3$disney_tv <- gsub("0", "", tv_df_3$disney_tv)

tv_df_3$all_platforms_tv <- paste(tv_df_3$netflix_tv, tv_df_3$hulu_tv, tv_df_3$prime_video_tv, tv_df_3$disney_tv)
tv_df_3$all_platforms_tv <- trimws(tv_df_3$all_platforms_tv, which = c("right"))

tv_df_3$all_platforms_tv <-
  ifelse((nchar(tv_df_3$all_platforms_tv)) >= 20, "multi_platform_show", tv_df_3$all_platforms_tv )
```

#### Rotten Tomatoes category column 
We're going to add another column to each data frame, where each show or film falls into a specific rotten tomatoes category:  certified fresh, fresh, and rotten. The Rotten Tomatoes scoring system is as follows: certified fresh content scores 75 on the Tomatomerter or above, fresh scores between 74 and 60, and rotten content scores 59 or below. Finally, we'll relocate the column.

##### Movie data frame
```{r movie rotten tomatoes category column, results='hide'}
movie_df_3$tomato_category_movie <- movie_df_3$rotten_tomatoes_movie
glimpse(movie_df_3)
sapply(movie_df_3, class)
vec <- c(10)
movie_df_3[ , vec] <- apply(movie_df_3[ , vec, drop=F], 1,
                               function(x) as.character(as.numeric(x)))
glimpse(movie_df_3)
movie_df_3$tomato_category_movie <-
  ifelse((movie_df_3$tomato_category_movie) >= 75, "certified_fresh", movie_df_3$tomato_category_movie)
movie_df_3$tomato_category_movie <-
  ifelse((((movie_df_3$tomato_category_movie) < 75)&((movie_df_3$tomato_category_movie) > 59)), "fresh", movie_df_3$tomato_category_movie)
movie_df_3$tomato_category_movie <-
  ifelse((movie_df_3$tomato_category_movie) <=59, "rotten", movie_df_3$tomato_category_movie)
movie_df_3<- movie_df_3 %>% 
  relocate(tomato_category_movie, .after=rotten_tomatoes_movie)
```

##### TV data frame
```{r tv rotten tomatoes category column, results='hide'}
tv_df_3$tomato_category_tv <- tv_df_3$rotten_tomatoes_tv
glimpse(tv_df_3)
sapply(tv_df_3, class)
vec <- c(10)
tv_df_3[ , vec] <- apply(tv_df_3[ , vec, drop=F], 1,
                            function(x) as.character(as.numeric(x)))
tv_df_3$tomato_category_tv <-
  ifelse((tv_df_3$tomato_category_tv) >= 75, "certified_fresh", tv_df_3$tomato_category_tv)
tv_df_3$tomato_category_tv <-
  ifelse((((tv_df_3$tomato_category_tv) < 75)&((tv_df_3$tomato_category_tv) > 59)), "fresh", tv_df_3$tomato_category_tv)
tv_df_3$tomato_category_tv <-
  ifelse((tv_df_3$tomato_category_tv) <=59, "rotten", tv_df_3$tomato_category_tv)
tv_df_3<- tv_df_3 %>% 
  relocate(tomato_category_tv, .after=rotten_tomatoes_tv)
```


#### Outliers with boxplots
We will examine the release_year and rotten_tomatoes for outliers. First, we'll check for outliers in release_year, then rotten_tomatoes. We can start with a boxplot to check for outliers. Then we'll use summary and the interquartile range (IQR) to find extreme outliers and remove them. 

##### Movie data frame
```{r movie year outliers, results='hide'}
boxplot(movie_df_3$release_year_movie)
summary(movie_df_3$release_year_movie)
IQR(movie_df_3$release_year_movie)
Tmin = 2006-(3*12) 
Tmax = 2018+(3*12) 
movie_df_3$release_year_movie[which(movie_df_3$release_year_movie < Tmin | movie_df_3$release_year_movie > Tmax)]
```
We have 642 data points that are labeled as outliers here. We'll make a separate data frame that removes the outliers from the time line, but first we will examine the rotten_tomatoes column. 
```{r movie rotten tomatoes outliers}
boxplot(movie_df_3$rotten_tomatoes_movie)
summary(movie_df_3$rotten_tomatoes_movie)
IQR(movie_df_3$rotten_tomatoes_movie)
Tmin = 44- (3*18)
Tmax = 62+ (3*18)
```
The Tmin and Tmax for extreme outliers do not exceed the data in the data frame, so no outliers will be removed. We can now make a data frame for the movies data frame.
```{r movie outlier data frame}
movie_sans_outlier_df <- movie_df_3
movie_sans_outlier_df <- movie_sans_outlier_df %>% 
  filter(release_year_movie >= 1970)
```

##### TV data frame
We'll examine the TV data frame, both in release_year and rotten_tomatoes. Then we'll make a new data frame as needed. 
```{r tv year outliers, results='hide'}
boxplot(tv_df_3$release_year_tv)
summary(tv_df_3$release_year_tv)
IQR(tv_df_3$release_year_tv)
Tmin = 2011-(3*7) 
Tmax = 2018+(3*7) 
tv_df_3$release_year_tv[which(tv_df_3$release_year_tv < Tmin | tv_df_3$release_year_tv > Tmax)]
```
We have 207 data points that are labeled as outliers here. We'll make a separate data frame that removes the outliers from the time line, but first we'll examine the rotten_tomatoes column.
```{r tv rotten tomatoes outliers}
boxplot(tv_df_3$rotten_tomatoes_tv)
summary(tv_df_3$rotten_tomatoes_tv)
IQR(tv_df_3$rotten_tomatoes_tv)
Tmin = 36- (3*24)
Tmax = 60+ (3*24)
```
The Tmin and Tmax for extreme outliers do not exceed the data in the data frame, so no outliers will be removed. We can now make a data frame for the TV data frame.
```{r tv outliers data frame}
tv_sans_outlier_df <- tv_df_3
tv_sans_outlier_df <- tv_sans_outlier_df %>% 
  filter(release_year_tv >= 1990)
```

